---
title: "Tutorial"
author: "Tessera Crockett"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today we will be installing an R package that will help us process and prepare raw data to be analyzed. This data is resulting from an experiment on the wetland plant _Juncus effusus_, looking at the response to differing doses of ammonium. 
The goal of this package is to help us create useful biomass and height increase information based on this raw data, and then analyze the two to see if plant height increase can predict biomass increase. 
First lets install the package from github using the below code.

```{r, eval=FALSE}
devtools::install_github("tesseracrockett/crockett.pkg")
```   

Before we look at each function, make sure to load the package into your local or studio version of R.

```{r}
library(crockett.package)
```

## Determining root/shoot ratio for plant biomass

Alone, above_ground or belowground biomass are not particularly reflective of what a given plant may have experienced during it's growth season, however root to shoot ratio, a ratio (belowground/above_ground) that includes both pieces of plant growth information, can be very informative. Having a function to compute this ratio into a new column that can then be analyzed helps prevent errors typically found in hand calculations and provides this information quickly. 
The expected inputs are specific to the tibble (dataframe) "ecotox" and are the above_ground and belowground columns, which is biomass weight in grams. The expected output is a new column that has calculated the root to shoot ratio for each sample, the closer to 1, the closer the plant was to having equal roots to shoots.  

Now lets load in the data frame and run this function. 
```{r}
library(tidyverse)
ecotox <- read_csv("../inst/extdata/ecotox.csv")
calc_rootshoot_ratio(data = ecotox, BELOWGROUND = ecotox$BELOWGROUND, ABOVE_GROUND = ecotox$ABOVE_GROUND)

```

Next, lets makes sure to save our updated tibble and view it to confirm the presence of the new root/shoot ratio column, as we will need this later.


```{r}
ecotox <- calc_rootshoot_ratio(data = ecotox, BELOWGROUND = ecotox$BELOWGROUND, ABOVE_GROUND = ecotox$ABOVE_GROUND)
view(ecotox)
```

Let's take a look at the next function in this package.

## Determining total vertical growth in plant samples

Our raw data provides us with the initial and final height of the plant, however in order to better analyze our data and compare to other information such as biomass, we need to determine the difference in plant height from initial to final height measurements taken. Running this function helps prevent errors in hand calculations or calculations by rows. 
The expected inputs are specific to the tibble (dataframe) "ecotox" and include the "HEIGHT_I" and "HEIGHT_F" columns, which are both reported in centimeters. The expected output is a new column titled "fh_diff" which has calculated the difference in the intial and final height of the plant in centimeters. 

Now, let's first load our tibble back in and then processs our second function.

```{r}
ecotox <- read_csv("../inst/extdata/ecotox.csv")
calc_finalheight_diff(data = ecotox, HEIGHT_F = ecotox$HEIGHT_F, HEIGHT_I = ecotox$HEIGHT_I)
```

As with our first function, we want to make sure we overwrite our tibble with the updated version that includes our 2nd new column. 

```{r}
ecotox <-calc_finalheight_diff(data = ecotox, HEIGHT_F = ecotox$HEIGHT_F, HEIGHT_I = ecotox$HEIGHT_I)
view(ecotox)
```

Next, let's take a look at our final function.

## Generating a 


