---
title: "Tutorial"
author: "Tessera Crockett"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

How do you install r pkg? 
Syntax for installing pkgs looks like 

```{r, eval=FALSE}
devtools::install_github("tesseracrockett/crockett.pkg")
```   
next thing, remind people to load ppkg
```{r}
library(crockett.pkg)
```

## Determining root/shoot ratio for plant biomass

Alone, above_ground or belowground biomass are not particularly informative of what a given plant may have experienced during it's growth season, however root to shoot ratio, a ratio (belowground/above_ground) that includes both pieces of plant growth information. Having a function to compute this ratio into a new column that can then be analyzed helps prevent errors in hand calculations and provides this information quickly. The expected inputs are the above_ground and belowground biomass weights, in grams. The expected output is a new column that has calculated the root to shoot ratio for each sample. 

```{r}
library(tidyverse)
ecotox <- read_csv("../inst/extdata/ecotox.csv")
calc_rootshoot_ratio(ecotox)

```